\LARGE Untangling the link between traits, size and growth rate in plants

### 2016-08-05 ms/ms.tex
Allocation model and figures

Functional traits capture core differences in the strategies plants use to generate and invest surplus energy \citep{Wright-2004, Chave-2009, Westoby-2002}. Although most plants have the same basic physiological function and key resource requirements (carbon, nitrogen and water), species differ considerably in rates at which resources are acquired and invested into different tissues. During the last decade, trade-offs related to prominent traits -- such as leaf mass per leaf area, wood density, leaf nitrogen per area, vessel diameter, maximum  height, and seed size -- have been quantified and trait values exist for up to 10\% of the world's 250000 plant species \citep{Cornwell-2014}. Increasingly, researchers have been using traits as potential indicators of the growth strategy and ecology of a species.

While the influence of plant traits on elements of plant physiological function is increasingly understood, attempts at using traits to understand demographic rates however have only met with mixed success \citep{Wright-2010, Poorter-2008}. In seedlings, the trait leaf mass per leaf area -- the central element of the leaf economics spectrum \citep{Wright-2004} -- has been tightly related to relative growth rate in plant mass \citep{Lambers-1992, Wright-2000}. Leaf mass per leaf area and its correlate leaf lifespan have also been linked to height growth rate for small seedlings and saplings \citep{Reich-1992, Poorter-2006}. These early successes prompted researchers to search for similar relationships in large plants. However, the results showed that in trees leaf mass per leaf area was not correlated with diameter growth rate \citep{Poorter-2008, Wright-2010, Herault-2011}. Meanwhile, other traits such as wood density showed strong relationships to growth in large plants \citep{Wright-2010}, but only weak relationships in seedlings \citep{Castro-1998}. Yet, until recently, most trait growth relationships had been quantified for one or few traits and at a single life- stage.

The picture is now emerging that the effect of traits on plant growth may be  modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}. In a recent meta-analysis of more than  500 correlations recorded from 103 studies, \citet{Gibert-2016} showed that the sign of the correlation between some traits (such as leaf mass per leaf area and maximum height) and growth rate was modified by size, while for other traits (such as wood density and assimilation rate per leaf area) the sign of the correlation remained the same, irrespective of size. Importantly, the direction of the correlations and their shifts with size, supported predictions made based on understanding of trait-based trade-offs and their interaction with plant function \citep{Gibert-2016}.

The challenge of interpreting diverse empirical results seeking to link traits to growth rate has been all the harder because we currently lack clear expectations on the expected signal. Generating such expectations is one of the primary role for theory \citep{Kokko-2007}. Current empirical results suggest the effect of traits on growth changes with size and possibly also light environment, yet current theory says little about how such relationships might come about. A widely-used model shows that for seedlings, mass-based relative growth rate is linearly and negatively related to leaf mass per leaf area \citep{Lambers-1992, Cornelissen-1996, Wright-2000}. An extension of the model suggests a similar relationship may hold at larger sizes \citep{Enquist-2007}. But as described above, this prediction is not supported by empirical results. Meanwhile, theoretical predictions on how other traits should influence growth are largely absent..

A further problem with existing theory is that the effects of traits are all realised via influences on mass production \citep{Enquist-2007}, whereas the effect of traits such as leaf mass per leaf area and wood density is via allocation to different tissues \citep{Falster-2011, Gibert-2016}. A second issue with theory focussed solely on mass production is that measuring mass production is only practical for small plants that can be easily harvested. For larger plants, growth is more often measured as increment in either diameter or height. For example, the Centre for Tropical Science monitors diameter growth in 49 large permanent plots across forests worldwide \citep{Anderson-2015}; forestry agencies have likewise recorded growth in diameter in 1000's of plots across the globe \citep{Purves-2008, Kunstler-2015}. But as we outline further below, mass production is only one component of plant growth; meaning theory focussing solely on mass production will always be limited.

Ultimately, trait based understanding must b integrated within a more general model of plant growth, also able to capture patterns of growth through ontogeny. Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{Sillett-2010, King-2011} or diameter growth \citep{Herault-2011}.  In contrast, basal-area growth continues to increase with size \citep{Sillett-2010, Stephenson-2014}, due to an increasing influence of stem turnover. All growth measures decrease sharply with size when expressed as relative growth rates \citep{Rees-2010, Iida-2014}. Shade tolerance tends to decrease with increasing size \citep{Lusk-2006}, due to increasing costs of maintaining stems \citep{Givnish-1988}, and also vary with traits \citep{Lusk-2006,Poorter-2006}. These diverse phenomena -- summarised in Table \ref{tab:phenomena} -- are in need of a comprehensive and integrated explanation.

In this paper we show how a recently published mechanistic framework \plant{} \citep{Falster-2016} describing the potential effect of traits on growth throughout a plant's life-cycle can explain diverse phenomena (Table \ref{tab:phenomena}). The model employs the same framework for partitioning growth used by \citet{Gibert-2016} in their meta-analysis, building on and extending several related approaches \citep{Givnish-1988, Yokozawa-1995, Makela-1997, Moorcroft-2001, Falster-2011, King-2011} in which metrics such as height, diameter and mass growth are calculated as an emergent outcomes of several, independent, physiological processes. This approach is quite different to models derived from metabolic scaling theory \citep{Enquist-1999, Enquist-2007}, which attempt to organise everything around a single master ``scaling'' equation for mass growth, or statistically fitted growth models \citep[e.g.][]{Herault-2011, Ruger-2012, Iida-2014}. By disentangling the effects of plant size, light environment and traits on plant growth, we hope to provide a solid theoretical foundation for trait ecology and thus provide a platform for understanding growth across diverse species around the world.

@@

Our approach builds on the widespread approach used in many vegetation models of explicitly modelling the amounts in different tissues within the plant \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, DeKauwe-2014} (Fig. \ref{fig:conceptual}a). Here we consider the masses $M_i$, areas $A_i$, and diameters $D_i$ of tissues, with the subscript $i$ indicating: l=leaf, b=bark and phloem, s=sapwood, h=heartwood, r=root, a=live (l+b+s+r), t=total (l+b+s+h+r), and st=stem total (s+b+h). The total mass of living tissue is then $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_{\rm s}+M_{\rm r}$, and the standing mass of the plant is $M_{\rm t}=M_{\rm a}+M_{\rm h}$. A full list of variables, units and definitions is given in Table \ref{tab:params}.

We assume growth is fundamentally driven by energy income and subsequent distribution throughout the plant. Adopting a standard model, the amount of biomass available for growth, ${\rm d}B / {\rm d}t$ is given by the difference between income (total photosynthesis) and losses (respiration and turnover) within the plant \citep{Makela-1997, Thornley-2000, Falster-2011} (Fig. \ref{fig:conceptual}a, eqn \ref{eq:dbdt}). Photosynthesis is proportional to the average photosynthetic rate per unit leaf area $\bar{p}$ and total leaf area, $A_{\rm l} = M_{\rm l} / \phi$, where $\phi$ is leaf mass per leaf area. The constant $y$ in  eqn \ref{eq:dbdt} converts mol CO$_2$ into units of biomass. Although many models include a detailed physiological model for calculating $\bar{p}$ and $r_{\rm i}$, net biomass production eventually comes down to simple subtraction of income and loss.

\subsection{Decomposing of growth rate into components}

To model growth in either height, leaf area, basal area, standing mass or stem diameter requires that we account not just for net mass production, but also for the costs of building new tissues, allocation to reproduction, and architectural layout. Mathematically, these growth rates can be decomposed into a product of physiologically relevant terms \citep{Falster-2011, Gibert-2016} (Fig. \ref{fig:conceptual}b, eqns \ref{eq:dbdt}-\ref{eq:dDdt}). Importantly, most of these terms vary intrinsically with size. The insets in Fig. \ref{fig:conceptual}b show size-related patterns for a typical plant, from a specific allocation model derived below. However, it is important to recognise that eqns \ref{eq:dbdt}-\ref{eq:dDdt} are mathematically true, derived using the chain rule of differential calculus, and thus must hold true for any class of models where carbon allocation is important.

The first point to note is that the growth rate of all size metrics in eqns \ref{eq:dMtdt}-\ref{eq:dDdt} depend on the product of biomass production $\frac{{\rm d}B} {{\rm d}t}$ (from eqn \ref{eq:dbdt}) and the fraction of biomass allocated to growth of the plant, $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$. We assume the remaining fraction, $1- \frac{{\rm d}M_{\rm a}}{{\rm d}B}$, is allocated to reproduction. The term $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ starts high, presumably at 1.0 for seedlings, and then decreases through ontogeny, potentially to zero in fully mature plants. It is important to note here that $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ is the allocation of biomass \emph{after} replacing parts lost due to turnover. So even a plant with $\frac{{\rm d}M_{\rm a}}{{\rm d}B}=0$ will continue to produce some new leaves and increase in stem diameter, even if the net amount of live mass $M_{\rm a}$ does not increase any further.

@@

As noted above, most of eqns in Fig. \ref{fig:conceptual} arise as mathematical truisms. To make predictions from eqns \ref{eq:dbdt}-\ref{eq:dDdt}, however, requires a specific model of plant construction. The model described below can be considered a first-order functional-balance model, in that it seeks to maintain a balance between various elements of plant construction through ontogeny. Table \ref{tab:allometry} provides key equations of the model, while Table \ref{tab:params} provides estimates on key variable and fitted parameters.

Our model is founded on four key assumptions:

\noindent{\bf Assumption  1} -- There is an allometric (log-log) scaling relationship between the accumulated leaf area of a plant and its height. Relationship represented by eqn \ref{eq:ha}. Empirically motivated. Describes architectural strategy of plant.

\noindent{\bf Assumptions 2-4} -- An isometric relationship between the total leaf area and each of sapwood cross-sectional area (2), bark cross-sectional area (3), and total root mass (4). As per the pipe model, \citep{Shinozaki-1964} we assume that each unit of sapwood area supports a fixed area of leaf, so that the total canopy area of a plant relates to basal sapwood area eqn \ref{eq:as}. This leads to a solution for the total mass of sapwood in the plant as a function leaf area and height (eqn \ref{eq:ms2}). Bark and phloem tissue are modelled using an analogue of the pipe model (eqn \ref{eq:as}), leading to a similar equation as that for sapwood mass (eqn \ref{eq:msb}). Also consistent with pipe-model assumption, we assume a fixed ratio of root mass per unit leaf area (eqn \ref{eq:mr}).

\noindent{\bf Assumption 5} -- Vertical distribution of leaf within the canopy of a plant remains relatively constant through ontogeny. We follow the model of \citep{Yokozawa-1995} describing the vertical distribution of leaf area within the crowns of individual plants. This model can account for a variety of canopy profiles through a single parameter $\eta$. Setting $\eta=1$ results in a conical canopy, as seen in many conifers, while higher values, e.g. $\eta=12$ , give a top-weighted canopy profile similar to those seen among angiosperms.

Assuming lead immediately to an allometric model for plant size (Table \ref{tab:allometry}, see Supplementary material for derivations).

To first approximation, the assumptions, which are well-supported by data (Fig. \ref{fig:assumptions}).

Simple linear functions non-critical for results of this paper, explore using an example model, but any model with basic size effects will display same behaviour.

Substituting from Table \ref{tab:allometry} into eqns \ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives an explicit model for plant growth in relation to size.

@@

\subsection{Parameters}

\citep{Falster-2015b}: BAAD
\citep{Falster-2016}: \plant

@@

@@

This new growth model captures the intrinsically size-dependent nature of plant growth, recovering well-known empirical patterns \citep{Sillett-2010, King-2011}. Height growth shows a hump-shaped pattern with size, first increasing then decreasing. This pattern results from systematic changes in the four components of eqn \ref{eq:dhdt} with size (Fig. \ref{fig:conceptual}), including a strong decline in the fraction of plant that is leaf declines with increasing size (Fig. \ref{fig:conceptual}), and increasing reproductive allocation. In contrast, basal-area growth continues to increase with size \citep{Sillett-2010, Stephenson-2014}, due to an increasing influence of stem turnover. Diameter growth shows a weakly hump shaped curve \citep{Herault-2011}, tapering off slightly at larger sizes, in part because of allometric effects (eqn \ref{eq:dDdt}, but mainly because of increased reproductive allocation in older trees (Fig?). All growth measures decrease sharply with size when expressed as relative growth rates \citep{Iida-2014}.

@@

Unlike eqn \ref{eq:RGR}, eqs. \ref{eq:dhdt}, \ref{eq:dast} and \ref{eq:dDdt} predict a relationship between $\phi$ and growth that changes with plant size ( \ref{fig:lmA_growth_size}). Decreasing $\phi$ has two impacts on growth rate. First, lower $\phi$ increases marginal leaf deployment (${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$ by economising on construction costs. Second, lower $\phi$ decreases net production (${\rm d}P / {\rm d}t$), due to increased leaf turnover. Whether lower $\phi$ increases growth thus depends on the relative magnitude of these two effects. When plants are small the effect on leaf deployment rate is bigger and so decreasing $\phi$ increases growth rate. When plants are large, the influence of $\phi$ on leaf deployment rate is diminished, because the costs of building other supportive tissues (other terms in eqn \ref{eq:daldmt}) are larger (Fig. \ref{fig:conceptual}). The net result is that at larger sizes, low $\phi$ is no longer advantageous for growth (Fig. \ref{fig:lmA_growth_size}d).

If the trait of interest is leaf-construction cost, i.e. $x=\phi$, we can expand eqn \ref{eq:dg4} as follows. Note that $\frac{\partial \left(\frac{{\rm d}M_\textrm{t}} {{\rm d}A_{\rm l}}\right)}{\partial \phi} = 1$. Also, from eqn \ref{eq:dbdt} we obtain,
\begin{equation}\label{eq:dbdt2}
\frac{{\rm d}B}{{\rm d}t} = c_2 - A_{\rm l} A_4 \phi ^{1-B_4}
\end{equation}
where $c_2 = y ( A_{\rm l} \, (A - r_{\rm l}) + \sum_\textrm{i=b,s,r}{M_\textrm{i} \, r_\textrm{i}}) - (\sum_\textrm{i=b,s,r}{M_\textrm{i} \, hmax\textrm{i}})$is independent of $\phi$, and thus
\begin{equation}\label{eq:dbdt3}
\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial \phi}  =
(B_4-1) A_{\rm l} A_4\phi ^{-B_4}.
\end{equation}
eqn \ref{eq:dbdt3} shows that if $B4=1$, $\frac{ {\rm d}B} {{\rm d}t}$ is independent of $\phi$. In contrast, if $B4>1$ (which tends to be the case empirically), $\frac{ \textrm{d}P} {{\rm d}t}$ increases as $\phi$ increases, more acutely at low $\phi$.

Substituting into eqn \ref{eq:dg4}, we have the result that growth rate correlates negatively with traits if
\begin{equation} \label{eq:G6}
\frac{(B_4-1) A_{\rm l} A_4\phi ^{-B_4}}{c_2 - A_{\rm l} A_4 \phi ^{1-B_4}}
< \frac{1}{\phi
 + \frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_\textrm
 {b}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}}.
\end{equation}

@@

\subsubsection{Sapwood area per leaf area, $\theta$}

\subsubsection{Vessel size, $\kappa$}

\subsubsection{Root mass per leaf area, $\alpha_{\rm r1}$}

@@

>>>>

Functional traits capture core differences in the strategies plants use to generate and invest surplus energy \citep{Wright-2004, Chave-2009, Westoby-2002}. Although most plants have the same basic physiological function and key resource requirements (carbon, nitrogen and water), species differ considerably in rates at which resources are acquired and invested into different tissues. During the last decade, trade-offs related to prominent traits have been quantified and values for some traits -- such as leaf mass per leaf area, wood density, and seed size -- now exist for up to 10\% of the world's 250000 plant species \citep{Cornwell-2014}. Concomitantly, researchers have been using traits as potential indicators of the growth strategy and ecology of a species.

While the influence of plant traits on elements of plant physiological function has been increasingly quantified and understood, attempts at using traits to predict demographic rates have only met with mixed success \citep{Wright-2010, Poorter-2008}. In seedlings, the trait leaf mass per leaf area -- the central element of the leaf economics spectrum \citep{Wright-2004} -- was found to be tightly related to relative growth rate in plant mass \citep{Lambers-1992, Wright-2000}. Leaf mass per leaf area and its correlate leaf lifespan were also linked to height growth rate for small seedlings and saplings \citep{Reich-1992, Poorter-2006}. These early successes prompted researchers to search for similar relationships in large plants. However, the results showed that in trees leaf mass per leaf area was not correlated with diameter growth rate \citep{Poorter-2008, Wright-2010, Herault-2011}. Meanwhile, other traits such as wood density showed strong relationships to growth in large plants \citep{Wright-2010}, but less so in small plants \citep{Castro-1998}. In summary, traits did not correlate simply and consistently to growth rate.

The challenge of interpreting diverse empirical results seeking to link traits to growth rate is all the harder because while we lack clear theoretical expectations. Generating such expectations is one of the primary roles for theory \citep{Kokko-2007}. A widely-used model shows that for seedlings, mass-based relative growth rate may be linearly and negatively related to leaf mass per leaf area \citep{Lambers-1992, Cornelissen-1996, Wright-2000}. An extension of the model suggests a similar relationship may hold at larger sizes \citep{Enquist-2007}. But as described above, this prediction is not supported by empirical results. Meanwhile, theoretical predictions on how other traits should influence growth are largely absent. Without further guidance, many researchers have simply assumed traits to linearly map onto growth rates, i.e. there is a ``fast'' and ``slow'' growth end of the trait spectrum.

A further problem with existing theory is that the effects of traits are all realised via influences on mass production \citep{Enquist-2007}, whereas the influence of some traits such as leaf mass per leaf area and wood density is via allocation to different tissues rather than mass production \citep{Falster-2011, Gibert-2016}. A second issue with theory focussed solely on mass production is that measuring mass production is only practical for small plants that can be easily harvested. For larger plants, growth is more often measured as increment in either diameter or height. For example, natural research centres and forestry agencies have been recording stem diameter growth in 1000's of plots across the globe \citep{Purves-2008, Anderson-2015, Kunstler-2016}. As we outline further below, mass production is only one component of plant growth; meaning theory focussing solely on mass production will always be limited.

Ultimately, trait based understanding must be integrated within a more general model of plant growth, also able to capture patterns of growth through ontogeny. Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{Sillett-2010, King-2011} or diameter growth \citep{Herault-2011}. In contrast, basal-area growth continues to increase with size \citep{Sillett-2010, Stephenson-2014}, due to an increasing influence of stem turnover. And all growth measures decrease sharply with size when expressed as relative growth rates \citep{Rees-2010, Iida-2014}. Meanwhile shade tolerance tends to decrease with increasing size \citep{Lusk-2006}, due to increasing costs of maintaining stems \citep{Givnish-1988}, and also vary with traits \citep{Lusk-2006,Poorter-2006}. These diverse phenomena -- summarised in Table \ref{tab:phenomena} -- are deserving of a comprehensive and integrated explanation.

In this paper we show how a new mechanistic framework -- called {\plant} \citep{Falster-2016} -- offers fresh insight into the way traits influence plant growth throughout across the life-cycle and explain diverse phenomena (Table \ref{tab:phenomena}). Our primary aim is to explain a pattern that has been gradually emerging and recently solidified: that the effect of traits on plant growth may be modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}. A recent meta-analysis of 103 studies reporting \textgreater 500 correlations by \citet{Gibert-2016} provides the most compelling evidence. \citet{Gibert-2016} showed that the sign of the correlation between some traits (such as leaf mass per leaf area and maximum height) and growth rate was modified by size, while for other traits (such as wood density and assimilation rate per leaf area) the sign of the correlation remained the same, irrespective of size. Importantly, the direction of the correlations and their shifts with size supported predictions generated via a framework describing how traits influence growth. Here we extend the results of taken by \citet{Gibert-2016} to provide a complete, functioning model and explain, physiologically, why it makes sense that the effect of traits changes with size. By disentangling the effects of plant size, light environment and traits on plant growth, we hope to provide a solid theoretical foundation linking trait-based trade-offs to demography and thus provide a platform for understanding growth across diverse species around the world.

@@

We build on the widespread approach used in many vegetation models of explicitly modelling the amounts in different tissues within the plant \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, DeKauwe-2014} (Fig. \ref{fig:conceptual}a). Here we consider the masses $M_i$, areas $A_i$, and diameters $D_i$ of tissues, with the subscript $i$ indicating: l=leaf, b=bark and phloem, s=sapwood, h=heartwood, r=root, a=live (l+b+s+r), t=total (l+b+s+h+r), and st=stem total (s+b+h). The total mass of living tissue is then $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_{\rm s}+M_{\rm r}$, and the standing mass of the plant is $M_{\rm t}=M_{\rm a}+M_{\rm h}$. A full list of variables, units and definitions is given in Table \ref{tab:definitions} and Table \ref{tab:traits}, with further detail on the parameter values applied given in Table \ref{tab:params}.

We assume growth is fundamentally driven by energy income and subsequent distribution throughout the plant. According to a standard model, the amount of biomass available for growth, ${\rm d}B / {\rm d}t$ is given by the difference between income (photosynthesis) and losses (respiration and turnover) within the plant \citep{Makela-1997, Thornley-2000} (Fig. \ref{fig:conceptual}a, eqn \ref{eq:dbdt}). Photosynthesis is proportional to the average photosynthetic rate per unit leaf area $\bar{p}$ and total leaf area, $A_{\rm l} = M_{\rm l} / \phi$, where $\phi$ is leaf mass per leaf area. The constant $y$ in eqn \ref{eq:dbdt} converts mol CO$_2$ into units of biomass. We assume that  that $\bar{p}$ increases with light but with diminishing returns, and respiration and turnover rates of different tissues are constants that might differ by trait (see below). While many vegetation models use a detailed physiological model for calculating $\bar{p}$ and $r_{\rm i}$, e.g. as functions of temperature, such detail is not needed here. Even with detailed models, ${\rm d}B / {\rm d}t$ always comes down to simple subtraction of income and losses.

\subsection{Decomposing of growth rates into components}

To model growth in either height, leaf area, basal area, standing mass or stem diameter requires that we account not just for net mass production, but also for the costs of building new tissues, allocation to reproduction, and architectural layout. Mathematically, these growth rates can be decomposed into a product of physiologically relevant terms \citep{Falster-2011, Gibert-2016}. Relevant equations (eqns \ref{eq:dbdt}-\ref{eq:dDdt}) are given Fig. \ref{fig:conceptual}b. Importantly, most of these terms vary intrinsically with size. The insets in Fig. \ref{fig:conceptual}b show size-related patterns in different variables for a typical plant obtained from applying the allocation model derived below. However, it is important to recognise that eqns \ref{eq:dMtdt}-\ref{eq:dDdt} are mathematically true, derived using the chain rule of differential calculus, and thus must hold true for any class of models where carbon allocation is important.

Note that the growth rate of all size metrics in eqns \ref{eq:dMtdt}-\ref{eq:dDdt} depend on the product of biomass production $\frac{{\rm d}B} {{\rm d}t}$ (from eqn \ref{eq:dbdt}) and the fraction of biomass allocated to growth of the plant, $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$, which varies from 0-1. The remaining fraction,$1- \frac{{\rm d}M_{\rm a}}{{\rm d}B}$, is allocated to reproduction. $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ starts high, presumably at 1.0 for seedlings, and then decreases through ontogeny, potentially to zero in fully mature plants. It is important to realise that $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ is the allocation of biomass \emph{after} replacing parts lost due to turnover. So a plant with $\frac{{\rm d}M_{\rm a}}{{\rm d}B}=0$ will continue to produce some new leaves and increase in stem diameter, even while the net amount of live mass $M_{\rm a}$ does not increase any further.

@@

As noted above, most of eqns in Fig. \ref{fig:conceptual} arise as mathematical truisms. To make predictions from eqns \ref{eq:dbdt}-\ref{eq:dDdt}, however, requires a specific model of plant construction. The model described below can be considered a first-order functional-balance model. We might also call it ``isometric'', because the assumptions ensure the proportional relationships among the relevant variables are preserved as the plant grows. Table \ref{tab:allometry} provides key equations of the model (see Supplementary material \ref{sec:ff16} or \citet{Falster-2016} for full derivation). In particular, we assume:
\begin{description}
\item[Assumption 1:] A log-log scaling relationship with slope of 0.5 between the leaf area of a plant and its height (eqn \ref{eq:ha}).
\item[Assumptions 2:] A log-log scaling relationship with slope of 1.0 between the leaf area and cross-sectional area of sapwood in a plant (eqn \ref{eq:as}).
\item[Assumptions 3:] A log-log scaling relationship with slope of 1.0 between the leaf area and cross-sectional area of bark and phloem in a plant (eqn \ref{eq:ab}).
\item[Assumptions 4:] A log-log scaling relationship with slope of 1.0 between the leaf area and mass of fine roots in a plant (eqn \ref{eq:mr}).
\item[Assumption 5:] The vertical distribution of leaf area within the canopy of an individual plant remains relatively constant through ontogeny.
\end{description}

Assumption 1 accounts for the architectural layout of the plant. Assumptions 2-4 are implementations of the pipe model \citep{Shinozaki-1964}, whereby the area of conducting tissues are proportional to leaf area. To describe the vertical distribution of leaf area within the canopy of an individual plant, we follow the model of \citet{Yokozawa-1995}, which can account for a variety of canopy profiles through a single parameter $\eta_c$, varying from 0-1. Low values result in a conical canopy, as seen in many conifers, while higher values give a top-weighted canopy profile similar to those seen among angiosperms.

The functional-balance assumptions are well-supported by data, at least to first approximation. Fig. \ref{fig:assumptions} shows data sourced from the Biomass and Allometry Database (\textsc{baad}) \citep{Falster-2015b}, which includes records for 21084 individuals on 656 species for various size metrics of individual plants. Dashed lines indicate the slope of predictions under functional-balance assumptions in Table \ref{tab:allometry}a. As is evident, the lines fitted within each species roughly align with predicted functional-balance assumption.

Combined, these assumptions in Table \ref{tab:allometry}a lead directly to equations describing the mass of sapwood and bark in relation to leaf, and the amount of leaf in relation to height (Table \ref{tab:allometry}b). Substituting from Table \ref{tab:allometry} into eqns \ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives a the necessary components in Fig. \ref{fig:conceptual} and thereby an explicit model for plant growth in relation to size. Note also that no assumptions are made about the diameter and total mass of the stem in relation to other components, e.g. height: these relationships arise as emergent properties, via integrating stem turnover (eqns \ref{eq:ha}- \ref{eq:hm}).

The most important prediction arising from the eqns in Table \ref{tab:allometry} is that the amount of live tissue supporting each unit of leaf area increases with plant height. We can derive an explicit prediction in relation to the assumed allocation model:
\begin{equation}\label{eq:msml}
\frac{M_{\rm b} + M_{\rm s}}{A_{\rm l}} = (1+ \alpha_{\rm b1}) \,\theta \, \rho \, \eta_c \, H.
\end{equation}
Note that $\alpha_{\rm b1}, \theta, \rho, \eta_c$ are traits, i.e. properties that are assumed approximately constant through ontogeny, so that their product is a constant. Again, the predicted relationship is supported by available data from the \textsc{baad} (Fig. \ref{fig:support_costs}).

@@

Now consider how trait-based trade-offs can be embedded into the above growth model. It is essential that any trait include both a benefit and cost in terms of plant function and/or life history; otherwise we would expect never-ending directional trait values towards more beneficial values. For present purposes, we consider five prominent traits for which we can posit specific costs and benefits, outlined in Table \ref{tab:traits}. In postulating potential benefits and costs, we only consider direct biophysical consequences of varying a trait, i.e. we do not try to account for every single variable that may sometimes correlate with a given trait, only those for which there is thought to be an inevitable trade-off. The trade-offs described in Table \ref{tab:traits} are then implemented as follows.
\begin{description}
\item[Seed mass] Effects of the trait seed mass, denoted by $\omega$, are naturally embedded in the equations determining fecundity and size of seeds. We assume larger seeds result in larger seedlings, thus the energetic tradeoff arising from seed size translates into a demographic tradeoff between the number of seedlings produced by a parent and the initial size of seedlings.
\item[Height at maturation] ........
\item[Nitrogen per unit leaf area] We allow for the maximum photosynthetic capacity of the leaf to vary with leaf nitrogen per unit area $\nu$, as $A_{\rm max} = \beta_{{\rm lf1}} \, \left(\frac{\nu}{\nu_0}\right)^{\beta_{{\rm lf5}}}$, where $\beta_{{\rm lf1}}$, $\nu_0$ and $\beta_{{\rm lf5}}$ are constants. Respiration rates per unit leaf area are also assumed to vary linearly with leaf nitrogen per unit area, as $\beta_{{\rm lf4}}\, \nu$.
\item[Leaf mass per unit area] The trait leaf mass per unit area, denoted by $\phi$, directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$. In addition, we link $\phi$ to the rate of leaf turnover, based on a widely observed scaling relationship from \citet{Wright-2004}: $k_{\rm l} = \beta_{{\rm kl1}} \, \left(\frac{\phi}{\phi_0}\right)^{-\beta_{{\rm kl2}}}$ where $\beta_{{\rm kl1}}$, $\phi_0$ and $\beta_{{\rm kl2}}$ are empirical constants.
\item[Wood density] The trait wood density, denoted by $\rho$, directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$. The cost of cheaper wood is increased mortality rate, which is not included in the growth equations but itself an important demographic outcome. The rate of sapwood respiration per unit volume is assumed to be constant, so sapwood respiration per unit mass varies as $r_{\rm s}  = \frac{\beta_{{\rm rs1}}}{\rho},$ where $\beta_{{\rm rs1}}$ is a default rate per volume of sapwood. Similarly, the rate of bark respiration per unit mass varies as $r_{\rm b} = \frac{\beta_{{\rm rb1}}}{\rho},$ with $\beta_{{\rm rb1}} = 2 \beta_{{\rm rs1}}$.
 \end{description}

@@

The 
Parameters

\citep{Falster-2016}: \plant

@@

This new growth model captures the intrinsically size-dependent nature of plant growth, recovering well-known empirical patterns (Table \ref{tab:phenomena}). Height growth shows a hump-shaped pattern with size, first increasing then decreasing. This pattern results from systematic changes in the four components of eqn \ref{eq:dhdt} with size (Fig. \ref{fig:conceptual}), including a strong decline in the fraction of plant that is leaf declines with increasing size (Fig. \ref{fig:conceptual}), and increasing reproductive allocation. In contrast, basal-area growth continues to increase with size \citep{Sillett-2010, Stephenson-2014}, due to an increasing influence of stem turnover. Diameter growth shows a weakly hump shaped curve \citep{Herault-2011}, tapering off slightly at larger sizes, in part because of allometric effects (eqn \ref{eq:dDdt}, but mainly because of increased reproductive allocation in older trees (Fig?). All growth measures decrease sharply with size when expressed as relative growth rates \citep{Iida-2014}.

@@

Unlike eqn \ref{eq:RGR}, eqs. \ref{eq:dhdt}, \ref{eq:dast} and \ref{eq:dDdt} predict a relationship between $\phi$ and growth that changes with plant size ( \ref{fig:lmA_growth_size}). Decreasing $\phi$ has two impacts on growth rate. First, lower $\phi$ increases marginal leaf deployment ($\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ by economising on construction costs. Second, lower $\phi$ decreases net production (${\rm d}P / {\rm d}t$), due to increased leaf turnover. Whether lower $\phi$ increases growth thus depends on the relative magnitude of these two effects. When plants are small the effect on leaf deployment rate is bigger and so decreasing $\phi$ increases growth rate. When plants are large, the influence of $\phi$ on leaf deployment rate is diminished, because the costs of building other supportive tissues (other terms in eqn \ref{eq:daldmt}) are larger (Fig. \ref{fig:conceptual}). The net result is that at larger sizes, low $\phi$ is no longer advantageous for growth (Fig. \ref{fig:lmA_growth_size}d).

If the trait of interest is leaf-mass per unit leaf area, i.e. $x=\phi$, we can expand eqn \ref{eq:dg4} as follows. Note that $\frac{\partial \left(\frac{{\rm d}M_{\rm t}} {{\rm d}A_{\rm l}}\right)}{\partial \phi} = 1$. Also, from eqn \ref{eq:dbdt} we obtain,
\begin{equation}\label{eq:dbdt2}
\frac{{\rm d}B}{{\rm d}t} = c_2 - A_{\rm l} A_4 \phi ^{1-B_4}
\end{equation}
where $c_2 = y ( A_{\rm l} \, (A - r_{\rm l}) + \sum_{i={\rm b,s,r}}{M_{i} \, r_{i}}) - (\sum_{i={\rm b,s,r}}{M_{i} \, ???})$is independent of $\phi$, and thus
\begin{equation}\label{eq:dbdt3}
\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial \phi}  =
(B_4-1) A_{\rm l} A_4\phi ^{-B_4}.
\end{equation}
eqn \ref{eq:dbdt3} shows that if $B4=1$, $\frac{ {\rm d}B} {{\rm d}t}$ is independent of $\phi$. In contrast, if $B4>1$ (which tends to be the case empirically), $\frac{ {\rm d}P} {{\rm d}t}$ increases as $\phi$ increases, more acutely at low $\phi$.

Substituting into eqn \ref{eq:dg4}, we have the result that growth rate correlates negatively with traits if
\begin{equation} \label{eq:G6}
\frac{(B_4-1) A_{\rm l} A_4\phi ^{-B_4}}{c_2 - A_{\rm l} A_4 \phi ^{1-B_4}}
< \frac{1}{\phi
 + \frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}}.
\end{equation}

@@

@@

In what follows, the assumption of exact isometry is non-critical, what matters is that the amount of live biomass (i.e. excluding heartwood) needed to support an extra unit of leaf area increases with height (as Fig. \ref{fig:support_costs}).  Note that similar increases in construction and maintenance costs with plant size are embedded within various models of plant function \citep[e.g.][]{Givnish-1988, Makela-1997}.

The list of functional traits that are known to differ among plant species is already long and ever-increasing. However, far fewer have well established tradeoffs.
